### 2021-03-13 ms/data_readme.Rmd
Add Rmd to generate data README

---
output: rtf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE, cache = FALSE}
# TO CONVERT TO PLAIN TEXT:
# after rendering to rtf,
# 1. Copy the contents of the rtf file, paste to txt.
# 2. Replace all single line breaks with double line breaks
# (in Atom search-replace with grep: \n for \n\n)
# 3. Hard-wrap all text at 80 characters (in Atom with autoflow: cmd-alt-q)
# (optional: add name after first line)
# 4. Save as "README.txt"

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE,
  warning = FALSE, results = "hide", cache = FALSE)

# Load objects from targets workflow (all paths to data files)
tar_load(gameto_times_2012_file)
tar_load(sporo_dt_times_file)
tar_load(specimens_raw_file)
tar_load(filmy_dt_file)
tar_load(filmy_dt_chamber_file)
tar_load(filmy_habit_file)
tar_load(light_data_file)
```

This README.txt file was generated on `r format(Sys.time(), '%d %B, %Y')`

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\- GENERAL INFORMATION \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

Title of Dataset: Data from: Intergenerational niche differentiation in filmy
ferns (Hymenophyllaceae)

Author Information

Principal Investigator: Joel H. Nitta

Department of Integrated Biosciences, Graduate School of Frontier Sciences,
The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-8562, Japan
joelnitta@gmail.com

Associate or Co-investigators: James E. Watkins, N. Michele Holbrook,
Tristan W. Wang, Charles C. Davis

Date of data collection: 2012--2014

Geographic location of data collection: Moorea, French Polynesia

Information about funding sources or sponsorship that supported the
collection of the data: Funding provided in part by the National Science
Foundation (Doctoral Dissertation Improvement Grant DEB-1311169 to JHN and
CCD), Setup Funds from Harvard University to CCD, American Society of Plant
Taxonomists (Research Grant for Graduate Students to JHN), Garden Club of
America (Award in Tropical Botany to JHN), Harvard University Herbaria (Fernald
Fieldwork Fellowship to JHN), Society of Systematic Biologists (Graduate Student
Research Award to JHN), and Systematics Association (Systematics Research Fund
to JHN).

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

SHARING/ACCESS INFORMATION

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

Licenses/restrictions placed on the data, or limitations of reuse: CC0 1.0
Universal (CC0 1.0)

Recommended citation for the data: Nitta JH, Watkins JE, Holbrook NM, Wang TW,
Davis CC (2021) Data from: Intergenerational niche differentiation in filmy
ferns (Hymenophyllaceae) https://doi.org/10.6084/m9.figshare.14184572

Citation for and links to publications that cite or use the data: Nitta JH,
Watkins JE, Holbrook NM, Wang TW, Davis CC (2021) Intergenerational niche
differentiation in filmy ferns (Hymenophyllaceae) <<FIXME: ADD JOURNAL>>

Code for analyzing the data is available on github:
https://github.com/joelnitta/moorea_filmies

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

DATA & FILE OVERVIEW

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

File list (filenames, directory structure (for zipped files) and brief
description of all data files):

- 2012_gameto_dt_times.csv: Start and end times of desiccation tolerance (DT)
  test on fern gametophytes during the 2012 field season.
- 2012_sporo_dt_times.csv: Start and end times of desiccation tolerance (DT)
  test on fern sporophytes during the 2012 field season.
- data_raw.tar.gz: Zipped raw data files. For a description, see README.md in
  the zipped archive.
- fern_specimens.csv: Specimen data for ferns collected in French Polynesia
  primarily by Joel H. Nitta.
- filmy_dt.csv: Data from desiccation tolerance test on filmy ferns from
  from Moorea, French Polynesia.
- filmy_dt_chamber.csv: Temperature and humidity inside desiccation chambers during
  desiccation tolerance test.
- filmy_growth_habit.csv: Growth habit of filmy ferns from Moorea,
  French Polynesia.
- filmy_light_curves.csv: Light levels and chlorophyll fluorescence measured
  in filmy ferns from Moorea, French Polynesia with a chlorophyll
  fluorometer.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

METHODOLOGICAL INFORMATION

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

Description of methods used for collection/generation of data:

Desiccation tolerance (DT) and light responses were measured in sporophytes and
gametophytes of filmy ferns (family Hymenophyllaceae) on the island of Moorea,
French Polynesia.

Samples were collected in the field and stored in plastic bags with a small
amount of water to keep them fresh during transport to the lab.

To measure DT, pre-treatment maximum photochemical yield of photosystem II was
measured in fresh plants after a 10 min period of dark-adaptation using a
portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany). Samples were
then transferred to desiccation chambers containing saturated salts at three
different desiccation intensities or a control treatment with moist tissues
(100% RH), and water withheld for either a short (2 d) or long (15 d) interval.
Conditions inside the desiccation chambers were monitored during the experiment
using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging
every 10 min or Hobo ProV2 dataloggers logging every 5 min. Salts used for
desiccation and their corresponding mean water potentials and approximate
relative humidity and VPD are as follows: LiCl (-282 MPa, 18% RH, 2.45 kPa),
Mg(NO3)2 (-86 MPa, 58% RH, 1.25 kPa), and NaCl (-38 MPa, 80% RH, 0.60 kPa).
Following the desiccation treatment, plants were rewetted and yield of
photosystem II was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Eight individuals per treatment were used for sporophytes. No replication by
species was possible for gametophytes due to their cryptic morphology. All
gametophytes were subjected to the same treatment (2 d at -86 MPa). Gametophytes
were later identified to species by DNA barcoding.

Relative water content (RWC) was measured in a subset of samples by recording
the mass of samples prior to the DT test (fresh mass), at each step of the DT
test (turgid mass), then after drying them overnight in a drying oven at 65 °C
following the DT test (dry mass). Relative water content was calculated as RWC =
(TM - DM) / (FM - DM), where TM is turgid mass, DM is dry mass, and FM is fresh
mass. Relative water content was not calculated for gametophytes, as these were
too small to measure accurately with the balances available.

To measure light responses, rapid light response curves were constructed for
each species by measuring photosynthetic yield at gradually increasing levels of
photosynthetically active actinic light (400 nm to 700 nm) with the Light Curve
function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh,
Effeltrich, Germany) as described by the manufacturer.

Fieldwork was done under permits issued by the French Polynesian Government
(Délégation à la Recherche) and the Haut-commissariat de la République en
Polynésie Francaise (Protocole d’Accueil 2012–2014).

fern_specimens.csv, filmy_dt.csv, filmy_dt_chamber.csv, and
filmy_light_curves.csv were generated by processing the raw data files using
scripts available at https://github.com/joelnitta/moorea_filmies (scripts
beginning with "clean_" in the "R/" directory).

For full methods, see Nitta JH, Watkins JE, Holbrook NM, Wang TW, Davis CC (2021) <<FIXME: ADD JOURNAL>>

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

DATA-SPECIFIC INFORMATION

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r gameto-dt-times}
gameto_dt_times <- read_csv(here::here(gameto_times_2012_file))
```

2012_gameto_dt_times.csv: Start and end times of desiccation tolerance (DT)
experiment on fern gametophytes during the 2012 field season. All gametophytes
were desiccated 2 d at -86 MPa. For other field seasons (2013, 2014), start and
end times were recorded during each measurement, so are not included in this
file.

Number of variables: `r ncol(gameto_dt_times)`

Number of cases/rows: `r nrow(gameto_dt_times)`

Variable list:

- date_time: Date and time, formatted YYYY-MM-DD HH-MM-SS.
- event: Start or end of experiment.
- group: Code corresponding to site and growth habit of samples.
- cluster: Number assigned to each group of samples (gametophyte individuals)
that were included together in a single experiment.

Missing data codes: No missing data.

Specialized formats or other abbreviations used: For 'group', 'ter' indicates
terrestrial samples, 'epi' indicates epiphytic samples; number indicates site
as follows: 1 = Three Pines 200m, 2 = Mt. Tohiea 400m, 3 = Mt. Tohiea 600m,  4 =
Mt. Tohiea 800m, 5 = Mt. Tohiea 1000m, 6 = Mt. Tohiea 1170m. For a description
of sites, see Nitta, Joel H.; Meyer, Jean-Yves; Taputuarai, Ravahere; Davis,
Charles C. (2016), Data from: Life cycle matters: DNA barcoding reveals
contrasting community structure between fern sporophytes and gametophytes,
Dryad, Dataset, https://doi.org/10.5061/dryad.df59g

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r sporo-dt-times}
sporo_dt_times <- read_csv(here::here(sporo_dt_times_file))
```

2012_sporo_dt_times.csv: Start and end times of desiccation tolerance (DT) test
on fern sporophytes during the 2012 field season. For other field seasons
(2013, 2014), start and end times were recorded during each measurement, so are
not included in this file.

Number of variables: `r ncol(sporo_dt_times)`

Number of cases/rows: `r nrow(sporo_dt_times)`

Variable list:

- date_time: Date and time, formatted YYYY-MM-DD HH-MM-SS
- event: Start or end of experiment.
- salt: Name of salt used to maintain constant humidity in the desiccation
chamber (see methods above for corresponding relative humidity).
- dry_time: Length (d) of desiccation.
- species: Species included in experiment.

Missing data codes: No missing data.

Specialized formats or other abbreviations used: None.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

data_raw.tar.gz: Zipped raw data files. For a description, see README.md in
the zipped archive. These were used to generate the clean data files with code
available at https://github.com/joelnitta/moorea_filmies (scripts starting with
'clean_' in the 'R' folder).

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r specimens}
specimens <- read_csv(here::here(specimens_raw_file))
```

fern_specimens.csv: Specimen data for ferns collected in French Polynesia
primarily by Joel H. Nitta.

Number of variables: `r ncol(specimens)`

Number of cases/rows: `r nrow(specimens)`

Variable list:

- specimen_id: Unique key assigned to each specimen.
- specimen: Specimen voucher (collector name and number).
- collector: Name of primary collector.
- coll_num: Collection number.
- family: Family.
- genus: Genus.
- specific_epithet: Specific epithet.
- infraspecific_rank: Infraspecific rank.
- infraspecific_name: Infraspecific name.
- certainty: Taxonomic certainty.
- species: Genus and specific epithet.
- taxon: Genus, specific epithet, and any infraspecific names.
- country: Country.
- county: County (or approximate political division).
- locality: General locality.
- site: Name of collection site.
- elevation: Elevation (m).
- latitude: Latitude (decimal degrees).
- longitude: Longitude (decimal degrees).
- observations: Observations about the specimen.
- generation: Gametophyte or sporophyte.
- gameto_net_location: Location of gametophyte sampling grid.
- gameto_square: Grid square where gametophyte was sampled.
- gameto_habit: Gametophyte growth habit.
- date_collected: Collection date.
- other_collectors: Names of other collectors present.
- herbaria: Herbarium code where specimen is deposited.

Missing data codes: Missing data entered as 'NA'.

Specialized formats or other abbreviations used: Herbarium codes follow Index
Herbariorum (http://sweetgum.nybg.org/science/ih/)

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r filmy-dt}
filmy_dt <- read_csv(here::here(filmy_dt_file))
```

filmy_dt.csv: Data from desiccation tolerance test on filmy ferns from from
Moorea, French Polynesia.

Number of variables: `r ncol(filmy_dt)`

Number of cases/rows: `r nrow(filmy_dt)`

Variable list:

- species: Species.
- salt: Name of salt used to maintain constant humidity in the desiccation
chamber (see methods above for corresponding relative humidity).
- dry_time: Length (d) of desiccation.
- individual: Individual code assigned to each sample per species per treatment.
(sporophytes) or collection number (gametophytes).
- generation: Gametophyte or sporophyte.
- dataset: Name of corresponding raw dataset file (sporophytes only).
- yield_pre: Maximum photochemical yield of photosystem II (Fv/Fm) prior to
desiccation treatment.
- yield_30min: Fv/Fm 30 m after rewetting following desiccation treatment.
- yield_24hr: Fv/Fm 24 h after rewetting.
- yield_48hr: Fv/Fm 48 h after rewetting.
- yield_72hr: Fv/Fm 72 h after rewetting (gametophytes only).
- yield_dry: Fv/Fm in the desiccated state.
- weight_pre: Mass (g) prior to desiccation treatment.
- weight_desiccated: Mass (g) in the desiccated state.
- weight_30min: Mass (g) 30 m after rewetting.
- weight_24hr: Mass (g) 24 h after rewetting.
- weight_48hr: Mass (g) 48 h after rewetting.
- weight_dry: Mass (g) after drying overnight at 60 C.
- time_pre: Date and time just before starting desiccation treatment,
formatted YYYY-MM-DD HH-MM-SS.
- time_30min: Date and time 30 m after rewetting.
- time_24hr: Date and time 24 h after rewetting.
- time_48hr: Date and time 48 h after rewetting.
- time_72hr: Date and time 72 h after rewetting (gametophytes only).
- time_dry: Date and time when yield was measured in the desiccated state.

Missing data codes: Missing data entered as 'NA'.

Specialized formats or other abbreviations used: None.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r filmy-dt-chamber}
filmy_dt_chamber <- read_csv(here::here(filmy_dt_chamber_file))
```

filmy_dt_chamber.csv: Temperature and humidity inside desiccation chambers during
desiccation tolerance test, measured using Track-It RH/Temp dataloggers
(Monarch Instrument, Amherst, NH) logging every 10 min or Hobo ProV2 dataloggers
logging every 5 min.

Number of variables: `r ncol(filmy_dt_chamber)`

Number of cases/rows: `r nrow(filmy_dt_chamber)`

Variable list:

- date_time: Date and time, formatted YYYY-MM-DD HH-MM-SS.
- temp: Temperature (degrees Celsius)
- rh: Relative humidity (%)
- salt: Name of salt used to maintain constant humidity in the desiccation
chamber (see methods above for corresponding relative humidity).
- generation: Gametophyte or sporophyte, or "mixed" if samples of both generations
were present in the growth chamber.
- year: Year.
- species: Species.
- file: Name of file containing raw data.
- serial_no: Serial number of datalogger.

Missing data codes: Missing data entered as 'NA'.

Specialized formats or other abbreviations used: None.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r filmy-habit}
filmy_growth_habit <- read_csv(here::here(filmy_habit_file))
```

filmy_growth_habit.csv: Growth habit of filmy ferns from Moorea, French
Polynesia. Growth habit determined by field observation. Species were
categorized into one of four growth habits: terrestrial (growing in soil, with
true roots), saxicolous  (growing on rocks), low-elevation epiphyte (mostly
occurring outside of cloud  forest, i.e., below ca. 500 m), and high-elevation
epiphyte (mostly occurring  within cloud forest, i.e., above ca. 500 m).
Crepidomanes bipunctatum and Crepidomanes  humile were observed growing as both
epiphytes and saxicoles, but are coded as  epiphytes to distinguish them from
exclusively saxicolous species.

Number of variables: `r ncol(filmy_growth_habit)`

Number of cases/rows: `r nrow(filmy_growth_habit)`

Variable list:

- species: Species.
- habit: Growth habit.

Missing data codes: None.

Specialized formats or other abbreviations used: None.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r filmy-light-curves}
filmy_light_curves <- read_csv(here::here(light_data_file))
```

filmy_light_curves.csv: Light levels and chlorophyll fluorescence measured in
filmy ferns from Moorea, French Polynesia with a chlorophyll fluorometer.
Chlorophyll  fluorescence measured with "light curve" function of mini-PAM
chlorophyll fluorometer.

Number of variables: `r ncol(filmy_light_curves)`

Number of cases/rows: `r nrow(filmy_light_curves)`

Variable list:

- type: Type of fluorescence measurement, from mini-PAM raw data.
- no: Number of fluorescence measurement, from mini-PAM raw data (not unique, 
because this file includes measurements from multiple mini-PAM raw data files).
- f: Momentary fluorescence level before application of saturation pulse.
- fm: Maximum fluorescence level of the illuminated sample.
- par: Level of photosynthetically active radiation during fluorescence 
measurement.
- yield: Effective photochemical quantum yield of photosystem II.
- etr: Relative electron transport rate.
- date_time: Date and time of measurement, formatted YYYY-MM-DD HH-MM-SS.
- species: Species.
- individual: Individual code assigned to each sample per species per treatment.
- generation: Gametophyte or sporophyte.
- coll_num: J.H. Nitta collection number
- sporo_site: Name of site where sporophyte sample was collected.
- condition: Condition under which measurements were taken (lab or field).
- date: Date of measurement, formatted YYYY-MM-DD.
- outlier: "TRUE" or "FALSE", indicating whether measurement should be excluded 
from light curve analysis.
- light_id: Unique ID number corresponding to light curve measured from a single 
sample.

Missing data codes: Missing data entered as 'NA'.

Specialized formats or other abbreviations used: For a description of sporophyte
sites, see Nitta, Joel H.; Meyer, Jean-Yves; Taputuarai, Ravahere; Davis,
Charles C. (2016), Data from: Life cycle matters: DNA barcoding reveals
contrasting community structure between fern sporophytes and gametophytes,
Dryad, Dataset, https://doi.org/10.5061/dryad.df59g

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

CHANGE LOG

\-\-\-

2021-03-14

Generate this README file.